---
title: Machine Learning - Predictions with linear regressions
author: Michael Fuchs
date: '2019-01-11'
slug: machine-learning-predictions-with-linear-regressions
categories:
  - R
tags:
  - R Markdown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(caret)
library(car)
library(glmnet)
```

```{r results='hide', message=FALSE, warning=FALSE}
machine <- read_csv("machine.csv")
cars <- read_csv("cars.csv")
```


#Table of Content

+ 1 Introduction
+ 2 Presentation of the data records used
+ 3 Dividing the data into a training part and a test part
+ 4 Removing problematic features
+ 4.1 Machine dataset
+ 4.2 Cars dataset
+ 5 Assessing linear regression models
+ 6 How to check some summary outputs individually
+ 6.1 Residual analysis
+ 6.2 Significance tests for linear regression
+ 6.3 Residual standard error (RSE)
+ 6.4 R^2^
+ 7 Test set performance with the MSE 
+ 8 Problems with linear regression
+ 8.1 Multicollinearity
+ 8.2 Outliers
+ 8.2.1 Compare RSE
+ 8.2.2 Compare MSE
+ 9 Future selection
+ 9.1 Forward selection
+ 9.2 Backward selection
+ 9.3 Comparing the calculated MSE
+ 10 Regularization
+ 11 Conclusion


#1 Introduction

I have already written several general contributions on linear regression models (["here"](https://michael-fuchs.netlify.com/2018/09/21/regression-analysis/) and ["here"](https://michael-fuchs.netlify.com/2018/10/02/special-regression-analysis/)). Likewise, I have already taken up the issue of how linear regression models can be trained and tested (["Machine Learning - Training and Testing Sets: Regression Modeling"](https://michael-fuchs.netlify.com/2018/12/18/machine-learning-training-and-testing-sets-regression-modeling/)) and how unimportant variables can be identified (["Machine Learning - Regression Regularization"](https://michael-fuchs.netlify.com/2018/12/25/machine-learning-regression-regularization/)).

In this publication, the prediction of the dependent variables from two different data networks should be central. In particular, the potential for improvement of the predictive power of the created models shall be considered.


For this post the dataset *machine* from the UCI- Machine Learning Repository platform ["UCI"](https://archive.ics.uci.edu/ml/datasets.html) was used as well as the dataset *cars* from the caret-package. A copy of the records is available at  <https://drive.google.com/open?id=1tMtpU5xEkijF-GceVEmGwRQCVrhWNrvf> (machine) and <https://drive.google.com/open?id=1E_NWwWEBBby456SuHt3qxHrsDpTzcUuv> (cars).


#2 Presentation of the data records used

The first dataframe (machine) contains the characteristics of different CPU models, such as the cycle time and the amount cache memory. PRP will be the dependent variable.The variables that are superfluous for this study are excluded in advance.

```{r}
machine <- machine[, 4:10]
head(machine, n = 5)
```

The second datafame (cars) contains details about different used car components, such as the number of cylinders or the number of miles the car has been driven. 

```{r}
cars <- cars[, -1]
glimpse(cars)
```

#3 Dividing the data into a training part and a test part

```{r}
set.seed(223356)
machine_sampling_vector <- createDataPartition(machine$PRP, p = 0.85, list = FALSE)
machine_train <- machine[machine_sampling_vector,]
machine_train_labels <- machine$PRP[machine_sampling_vector]
machine_test <- machine[-machine_sampling_vector,]
machine_test_labels <- machine$PRP[-machine_sampling_vector]
machine_train_features <- machine[, 1:6]
```

```{r}
cars_sampling_vector <- createDataPartition(cars$Price, p = 0.85, list = FALSE)
cars_train <- cars[cars_sampling_vector,]
cars_train_labels <- cars$Price[cars_sampling_vector]
cars_test <- cars[-cars_sampling_vector,]
cars_test_labels <- cars$Price[-cars_sampling_vector]
cars_train_features <- cars[,-1]
```



#4 Removing problematic features

Highly correlated variables can influence the prediction of a linear model. From this point of view, it is recommended to overcome this in the first step and to exclude problematic features.

#4.1 Machine dataset

```{r}
machine_correlations <- cor(machine_train_features)
findCorrelation(machine_correlations)
```

As we can see, there is no correlation above 0.9 (by default). Let's check correlations of 0.75 by this. 
```{r}
findCorrelation(machine_correlations, cutoff = 0.75)
cor(machine_train$MMIN, machine_train$MMAX)
```

Ok we have three of them but 0.75 should not be problematic. 

```{r}
findLinearCombos(machine_correlations)
```


#4.2 Cars dataset

Let's have a look at the cars dataset.


```{r}
cars_cor <- cor(cars_train_features)
findCorrelation(cars_cor)
```

```{r}
findCorrelation(cars_cor, cutoff = 0.75)
```

Again three correlations between 0.75 and 0.9 (by default). One example of these three are the correlation betwenn the variables *Doors* and *Coupe*.


```{r}
cor(cars$Doors,cars$coupe)
table(cars$coupe,cars$Doors)
```

Another usefull function is the *findLinearCombos* function to detect exact linear combinations of other features.

```{r}
findLinearCombos(cars)
```

Here we are advised to drop the *coupe* and *wagon* columns, which are the 15th and 18th features, respectively, because they are exact linear combinations of other features. Note that the division into a training part and a test part must be redone.

```{r}
set.seed(323356)
cars <- cars[,c(-15, -18)]
cars_sampling_vector <- createDataPartition(cars$Price, p = 0.85, list = FALSE)
cars_train <- cars[cars_sampling_vector,]
cars_train_labels <- cars$Price[cars_sampling_vector]
cars_test <- cars[-cars_sampling_vector,]
cars_test_labels <- cars$Price[-cars_sampling_vector]
cars_train_features <- cars[,-1]
```


#5 Assessing linear regression models

Once all the problematic features have been identified and excluded, it is time to create the regression models.

```{r}
machine_model1 <- lm(PRP ~ ., data = machine_train)
cars_model1 <- lm(Price ~ ., data = cars_train)
```

Here the summary for the machine_model1:
```{r}
summary(machine_model1)
```


And here the summary for the cars_model1:
```{r}
summary(cars_model1)
```

Please note the following message from the output: "Coefficients: (1 not defined because of singularities)". This occurs because we still have features whose effect on the output is indiscernible form other features due to underlying dependencies. We can identify this variable using the *alias* function. 


```{r}
alias(cars_model1)
```

Now we know, that we have to remove the *Saturn* variable.

```{r}
cars_model2 <- lm(Price ~. -Saturn, data = cars_train)
summary(cars_model2)
```

Perfect!


#6 How to check some summary outputs individually

All the information which we receive with the summary command can also be calculated automatically:

#6.1 Residual analysis

A residual is simply the error our model makes for a paritcular observation. 

```{r}
summary(cars_model2$residuals)
mean(cars_train$Price)
```

With this information we can say that the average selling price of a car in our training data is around 21k Dollar, and 50% of our predictions are roughly within +- 1.6k of the correct value. 


A useful way to graphically compare the quantiles of the distributions of two quantitative variables is the quantile-quantile diagram.

```{r}
par(mfrow = c(2, 1))
machine_residuals <- machine_model1$residuals
qqnorm(machine_residuals, main = "Normal Q-Q Plot for CPU data set")
qqline(machine_residuals)
cars_residuals <- cars_model2$residuals
qqnorm(cars_residuals, main = "Normal Q-Q Plot for Cars data set")
qqline(cars_residuals)
```

As we can see the residuals from both models seem to lie reasonably close to the theoretical quantiles of a normal distribution, although the fit isn't perfect, as is typical with most real-world data.



#6.2 Significance tests for linear regression

Let's have a look at the machine_model1 output again: 

```{r}
summary(machine_model1)
```

We can compute some figures manually, eg. the t-value and p-value of the MYCT variable:


```{r}
q <- 3.940e-02 / 1.843e-02
q

pt(q, df = 172, lower.tail = F) * 2
```

#6.3 Residual Standard Error (RSE)

We define a metric known as the Residual Standard Error, which estimates the standard deviation of our model compared to the target function. 

We can compute the RSE for our two models using  the preceding formular as follows:

```{r}
n_machine <- nrow(machine_train)
k_machine <- length(machine_model1$coefficients) - 1
sqrt(sum(machine_model1$residuals ^ 2) / (n_machine - k_machine - 1))
```



```{r}
n_cars <- nrow(cars_train)
k_cars <- length(cars_model2$coefficients) - 1
sqrt(sum(cars_model2$residuals ^ 2) / (n_cars - k_cars - 1))
```


To interpret the RSE values for our two models, we neet to compare them with the mean of our output variables.

```{r}
mean(machine_train$PRP)

mean(cars_train$Price)
```


#6.4 R^2^

Now let's compute R^2^ manually to compare different regression models.

```{r}
compute_rsquared <- function(x, y) {
  rss <- sum((x - y) ^ 2)
  tss <- sum((y - mean(y)) ^ 2)
  return(1 - (rss / tss))
}

compute_rsquared(machine_model1$fitted.values, machine_train$PRP)
```
```{r}
compute_rsquared(cars_model2$fitted.values, cars_train$Price)
```


#6.5 Adjusted R^2^

We can also do the same manual calculation for the adjusted R^2^.

```{r}
compute_adjusted_rsquared <- function(x, y, k) {
  n <- length(y)
  r2 <- compute_rsquared(x, y)
  return(1 - ((1 - r2) * (n - 1) / (n - k - 1)))
}

compute_adjusted_rsquared(machine_model1$fitted.values, machine_train$PRP, k_machine)
```
```{r}
compute_adjusted_rsquared(cars_model2$fitted.values, cars_train$Price, k_cars)
```


#7 Test set performance with the MSE

Now it's time to check the test set performance of our models. We can do this by the Mean Squared Error (**MSE**).

```{r}
machine_model1_predictions <- predict(machine_model1, machine_test)
cars_model2_predictions <- predict(cars_model2, cars_test)
```

Here is the required function for it:
```{r}
compute_mse <- function(predictions, actual) { 
  mean( (predictions - actual) ^ 2 ) 
}
```

#7.1 MSE for the machine dataset

Now we can compare the training and the test MSE for both models

```{r}
compute_mse(machine_model1$fitted.values, machine_train$PRP)
```
```{r}
compute_mse(machine_model1_predictions, machine_test$PRP)
```

#7.2 MSE for the cars dataset

```{r}
compute_mse(cars_model2$fitted.values, cars_train$Price)
```
```{r}
compute_mse(cars_model2_predictions, cars_test$Price)
```


#8 Problems with linear regression

In my post ["Regression Analysis"](https://michael-fuchs.netlify.com/2018/09/21/regression-analysis/) paragraph 5 Model assumption I have already addressed several aspects that must be taken into account when creating a regression model. For the present investigation of the two datasets two aspects are dealt with in detail.


#8.1 Multicollinearity

Multicollinearity is a problem of regression analysis and occurs when two or more explanatory variables have a very strong correlation with each other. We can check this with the *vif* function.


```{r}
vif(cars_model2)
```

These values can also be calculated manually by us. Here is the example for the calculation of the value for the variable Sedan

```{r}
sedan_model <- lm(sedan ~ .-Price -Saturn, data = cars_train)
sedan_r2 <- compute_rsquared(sedan_model$fitted.values, cars_train$sedan)
1 / (1-sedan_r2)
```


#8.2 Outliers

Outliers should always be taken into account in predictive models, as they can greatly influence the forecast, as the following example shows:


```{r}
scatterplot(machine_model1$fitted.values, machine_model1$residuals, id = TRUE, xlab = "Fitted Values", ylab = "Residuals", main = "Searching for outliers")
```


As we can see in the graphic line 171 is an outlier. Now we check what effect it has when we exclude line 171 from the training part.
```{r}
machine_model2 <- lm(PRP ~ ., data = machine_train[!(rownames(machine_train)) %in% c(171),])

summary(machine_model2)
```

#8.2.1 Compare RSE

```{r}
sqrt(sum(machine_model1$residuals ^ 2) / (n_machine - k_machine - 1))
```
```{r}
k_machine <- length(machine_model2$coefficients) - 1
sqrt(sum(machine_model2$residuals ^ 2) / (n_machine - k_machine - 1))
```



#8.2.2 Compare MSE

```{r}
compute_mse(machine_model1_predictions, machine_test$PRP)
```



```{r}
machine_model2_predictions <- predict(machine_model2, machine_test)
compute_mse(machine_model2_predictions, machine_test$PRP)
```



#9 Future selection

Future selection is a machine learning approach, that uses only a subset of the available features for a learning algorithm. To do this we need the null-modell of both regression models.

```{r}
machine_model_null <- lm(PRP ~ 1, data = machine_train[!(rownames(machine_train)) %in% c(171),])
cars_model_null <- lm(Price ~ 1, data = cars_train)
```


There are two types of future selection to operate: the forward selection and the backward selection (both demonstrated below). Note that the lower the AIC, the better the model.

#9.1 Forward selection

```{r}
machine_model3 <- step(machine_model_null, scope = list(lower = machine_model_null, upper = machine_model2), direction = "forward")
```

#9.2 Backward selection

```{r}
cars_model3 <- step(cars_model2, scope = list(lower=cars_model_null, upper=cars_model2), direction = "backward")
```


#9.3 Comparing the calculated MSE

Now we can again calculate the MSE for the best models selected by the feature selection:

```{r}
machine_model3_predictions <- predict(machine_model3, machine_test)
compute_mse(machine_model3_predictions, machine_test$PRP)
```

```{r}
cars_model3_predictions <- predict(cars_model3, cars_test)
compute_mse(cars_model3_predictions, cars_test$Price)
```



#10 Regularization
 
In addition to the future selection just shown, a regularization can also be used. In the following two methods will be shown on the basis of the car dataset: the ridge method and the lasso method.
 
 
 
```{r}
cars_train_mat <- model.matrix(Price ~ .-Saturn, cars_train)[,-1]
lambdas <- 10 ^ seq(8, -4, length = 250)
cars_models_ridge <- glmnet(cars_train_mat, cars_train$Price, alpha = 0, lambda = lambdas)
cars_models_lasso <- glmnet(cars_train_mat, cars_train$Price, alpha = 1, lambda = lambdas)
```



Here we can see how the Ridge Regression and the Lasso method works:
 
```{r} 
layout(matrix(c(1, 2), 1, 2))
plot(cars_models_ridge, xvar = "lambda", main = "Ridge Regression")
plot(cars_models_lasso, xvar = "lambda", main = "Lasso")
```

Here is a calculation to determine the best lambda for the two methods:


```{r}
ridge.cv <- cv.glmnet(cars_train_mat, cars_train$Price, alpha = 0)
lambda_ridge <- ridge.cv$lambda.min
lambda_ridge
```
```{r}
lasso.cv <- cv.glmnet(cars_train_mat, cars_train$Price, alpha = 1)
lambda_lasso <- lasso.cv$lambda.min
lambda_lasso
```


Again we'll compare the new MSE: 

```{r}
cars_test_mat <- model.matrix(Price ~ . -Saturn, cars_test)[,-1]
cars_ridge_predictions <- predict(cars_models_ridge, s = lambda_ridge, newx = cars_test_mat)
compute_mse(cars_ridge_predictions, cars_test$Price)
```
```{r}
cars_lasso_predictions <- predict(cars_models_lasso, s = lambda_lasso, newx = cars_test_mat)
compute_mse(cars_lasso_predictions, cars_test$Price)
```




#11 Conclusion

We have seen how to significantly improve the performance of linear models.







